Local to global N input modelling

### 2025-06-09 docs/manuscript.qmd
Forgot about jekyll's underscore vendetta
Fix up macronutrient summary

library(magrittr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ggh4x)
library(here)
library(ggpubr)
library(patchwork)
library(qs)
library(arrow)
library(targets)
library(sf)
library(terra)
library(rnaturalearth)
library(rnaturalearthdata)
library(rnaturalearthhires)
library(dtplyr)
library(furrr)
library(RColorBrewer)
library(purrr)
library(future)
library(readxl)
library(units)
library(conflicted)
library(lubridate)
library(RColorBrewer)
conflicts_prefer(dplyr::select(), dplyr::filter(), dplyr::intersect(), .quiet = T)
# This markdown uses TinyTex - install with tinytex::install_tinytex()

install_unit("gfish")
install_unit("kgfish", "1000 gfish")
install_unit("tfish", "1000 kgfish")
```

@@

species_params <- qread(species_params_file)
pop_params <- qread(pop_params_file)
```

# Introduction

[This intro is copied verbatim from work done during one of the early workshops.]{.aside}

Aquaculture is now the dominant form of aquatic animal food (herein ‘seafood’) production and is expected to be the primary way we meet future seafood demand. Freshwater systems will likely continue to provide the majority of farmed seafood but marine aquaculture is also poised to expand substantially in numerous areas. Farmed marine fish and invertebrates are produced near exclusively in coastal waters, and nearly three quarters of this production is dependent on human-made feeds. Nearshore locations and feed inputs are necessary to maintain profitable and productive farming operations but coastal aquaculture generates a number of challenges. In the crowded coastal zone, aquaculture operations can conflict with other stakeholder uses such as recreation, fishing, renewable energy, transport, and tourism. 
And while farming marine fish typically generates a far smaller nutrient footprint than livestock farming, the overt nature of aquaculture in nearshore regions and evidence of  localised nutrient impacts around fish farms remains a primary public and scientific concern. Identifying strategies that reduce ecosystem impacts from fish farm waste therefore represents an important goal for improving marine aquaculture sustainability and maintaining the sector’s social licence to operate.

Aquaculture feeds represent an important lever for reducing nutrient waste impacts around fish farms. Like all farmed animals, fish and invertebrates must digest the nutrients contained in feeds before they can be used for growth. Any nutrients left undigested are egested as solid waste, and dissolved wastes are excreted as metabolic waste products. Further, some feed inevitably remains uneaten and is lost to the surrounding ecosystem. Particulate organic matter (both feed and faeces) that settles can simplify benthic communities as the oxygen demand from its decomposition drives the production of sulphides that kill less mobile faunal, encouraging a lower diversity of opportunistic scavengers and the growth of bacterial mats (e.g., Beggiatoa spp). Thus, the chemical composition of the ingredients used in aquaculture feeds and their digestibility for the farmed species has significant implications for the nature and reactivity of the waste generate by marine aquaculture.

Firstly the overall volume of nutrient waste is dictated by the nature and intensity of production, that is the farm size, the density of farmed animals and the feed requirements and efficiency of the species grown. SecondlDeposition of waste is heavily influenced by water depth and current speed at the farming site. Once

As farmed fish and invertebrates are fed, whatever

Nutand its impact on marine ecosystems is influenced by many factors. 
* Farm size
* Depth
* Current speed
* Benthic impact - sediment type/faunal assemblages/wider marine community
* High turnover environments - nitrogen enriched areas
* Feed influences all of these things

The primary source of organic waste from fed aquaculture production comes from the excretion and faeces of the farmed animals and through uneaten feed that dissolves in the water column or settles on the benthos. The nature and impact of this waste are influenced heavily by the composition of the feeds fed to farmed animals.

## P2

* Waste from aquaculture farms and it’s impact is influenced by many things but the composition of feeds plays a central role.
* Waste from aquaculture farms has multiple sources. 
* The primary source of organic waste comes from the faeces and excretion of the fish or invertebrates.
* Uneaten feed produced another key source.
* The nature and impact of this waste are influenced heavily by the composition of the feeds fed to farmed animals
* Many marine fish are naturally carnivorous so diets used to be high in fishmeal and oil but increasing fishmeal and oil prices along with concerns over the sustainability of marine ingredients have led to a reduction in their use across multiple farmed taxa
* In lieu of fishmeal and oil, many plant-based ingredients such as soy protein concentrate, canola oil, and wheat gluten have replaced them. 
* Changes in feed composition influences the digestibility of the nutrients held in each feed and can alter the composition of waste.
* Of particular concern are changes (increases) to the presence of reactive nitrogen and phosphorus in coastal waters that could have an effect on eutrophication.

## P3

* Whether or not nutrients lead to eutrophication depends on the sensitivity of the receiving environment
* Ecosystems that are already enriched through natural processes and whose biota is well adapted to substantial fluxes in available nutrients (e.g. upwelling zones, dynamic coastal communities) may be less sensitive while oligotrophic ecosystem are likely to see considerable changes under nutrient enrichment scenarios.
* To understand the impact of aquaculture waste under present day or future scenarios we need to quantify the volume, nature, and location of mariculture waste and determine the sensitivity of the receiving environments to that waste. Yet only recent estimates even give us the estimated location of marine farms let alone the volume of nature of the waste produced.
To address this gap, we use existing maps of mariculture location with a bioenergtic model

# Methods
## Model and approach

[Unfinished. Need to work out how much should be transposed from @baldan_r_2018. One thing that definitely needs to be stated is the individual ingredient (rather than whole food) tracking, the use of cohorts, and the new method for calculating population.]{.aside}

We adapted the methods of @baldan_r_2018 to create a bioenergetic model that simulates individual growth and farm-scale production for Atlantic salmon and the resultant nutrient waste in the form of excess labile nitrogen and phosphorus. 
The model simulates growth at an individual level, calculating the change in individual weight through time using:

$$
\frac{dw}{dt} = \frac{A-C}{\epsilon}
$$

Where $w=$ is wet weight (t), $t=$ time (d), $A=$ anabolic rate (J t$^{-1}$), $C=$ the catabolic rate (J t$^{-1}$), $\epsilon=$ energy density of body tissues (J t$^{-1}$). 
Individual models were then upscaled using monte-carlo simulations to simulate size structure in a population. Size differences were achieved through different initial starting weights and ingestion rates for different finfish species. All individuals have a fixed mortality rate to simulate stocking and harvesting.

@@

worldmap <- ne_countries(scale = "medium", returnclass = "sf")

p1 <- ggplot(data = worldmap, aes(geometry = geometry)) +
  geom_sf(fill = "white", color = "dimgray") +
  geom_sf(data = farm_temp_means, aes(colour = mean_temp), size = 1.5) +
  coord_sf() +
  theme_classic()

p2 <- farm_temp_means %>% 
  ggplot(aes(y = mean_temp, x = abs(lat), colour = country)) +
  geom_point(size = 2) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = expression("Absolute latitude ("*degree*")"), y = expression("Mean temperature ("*degree*"C)")) +
  theme_classic() +
  theme(legend.position = "top", legend.title = element_blank())

p3 <- farm_temp_means %>% 
  mutate(hemisphere = case_when(lat < 0 ~ "S", T ~ "N")) %>% 
  ggplot(aes(x = mean_temp, fill = hemisphere)) +
  geom_histogram(position = "identity", colour = "black", alpha = 0.5, binwidth = 0.5) +
  labs(x = expression("Mean temperature ("*degree*"C)"), y = "Frequency") +
  theme_classic() +
  theme(legend.position = "top", legend.title = element_blank())

@@

```{r patchwork map}
#| fig-cap: |
#|   Possibly use a patchwork map to show spatial results?
#| fig-height: 9
#| fig-width: 9.5

worldmap <- ne_countries(scale = "large", returnclass = "sf")

AUS_bounds <- c(144, 149.5, -39.75, -44)
CAN1_bounds <- c(-132, -122, 47.75, 54.25)
CAN2_bounds <- c(-70, -54, 43, 48.5)
CHI_bounds <- c(-77.5, -62.5, -56, -25)
EUR_bounds <- c(-26, 30, 51, 71)

p_ALL <- ggplot(data = worldmap, aes(geometry = geometry)) +
  geom_sf(fill = "white", color = "dimgray") +
  geom_sf(data = farm_temp_means, aes(colour = mean_temp), size = 1.5) +
  coord_sf() +
  theme_classic()

no_margins <- function() {
  theme(legend.position = "none", 
        plot.margin = margin(),
        axis.title = element_blank())
}

p_Aus <- p_ALL + coord_sf(xlim = AUS_bounds[1:2], ylim = AUS_bounds[3:4], expand = FALSE) + no_margins()
p_Eur <- p_ALL + coord_sf(xlim = EUR_bounds[1:2], ylim = EUR_bounds[3:4], expand = FALSE) + no_margins()
p_Chi <- p_ALL + coord_sf(xlim = CHI_bounds[1:2], ylim = CHI_bounds[3:4], expand = FALSE) + no_margins()
p_Can <- p_ALL + coord_sf(xlim = CAN1_bounds[1:2], ylim = CAN1_bounds[3:4], expand = FALSE) + no_margins()
p_USA <- p_ALL + coord_sf(xlim = CAN2_bounds[1:2], ylim = CAN2_bounds[3:4], expand = FALSE) + no_margins()

subplot <- (p_Eur/p_Aus) + plot_layout(heights = c(1.5, 1))
finalplot <- (p_Can + p_USA) / (p_Chi + subplot) + 
  plot_layout(widths = c(1, 1.5), heights = c(1, 2)) &
  theme(plot.margin = margin(0, 0, 0, 0))
finalplot
```

```{r fig-cohorts}
#| fig-cap: Example of cohorts
#| eval: false

fnms <- output_cohort_growth_data_path %>% 
  list.files(full.names = T) %>% 
  str_subset("biomass")

df <- fnms[1] %>% read_parquet() %>% 
  filter(feed == "reference") %>% 
  mutate(mean = set_units(mean, "g") %>% set_units("t"),
         sd = set_units(sd, "g") %>% set_units("t"))

ggplot(df, aes(x = yday, y = mean, ymin = mean-sd, ymax = mean+sd, colour = as.factor(cohort))) +
  geom_line(linewidth = 0.75) +
  geom_ribbon(linetype = "dotted", alpha = 0) +
  scale_x_continuous(breaks = seq(0, 2000, 200)) +
  labs(y = "Farm biomass", x = "Day of the year") +
  theme_classic() +
  theme(legend.position = "none", aspect.ratio = 0.75)

yrs <- max(df$yday)/365
harvests <- df %>% 
  group_by(cohort) %>% 
  reframe(mean = maxna(mean),
          sd = maxna(sd))

sum(harvests$mean)/yrs
mean(harvests$sd)/yrs
```

### Feeding rate

df %>% 
  ggplot(aes(x = water_temp, y = ing_pot)) +
  geom_line(linewidth = 0.75) +
  geom_line(aes(y = food_prov), linewidth = 0.75, linetype = "dotted") +
  scale_y_continuous(breaks = seq(0,1.5,0.25)) +
  scale_x_continuous(breaks = seq(-5,30,5), limits = c(-5,26)) +
  labs(x = expression("Water temperature ("*degree*"C)"), y = "Relative food ingested/provided") +
  theme_classic() +
  theme(text = element_text(colour = "black", family = "serif", size = 12))
```

@@

Parameter sensitivity was tested by varying each parameter in @tbl-params by $\pm 10$% and calculating sensitivity as:

$$
S_p = \frac{(P_{1.1} - P_{0.9})}{0.2 \times P_1}
$$

where $S$ is the sensitivity of the measured variable to changes in parameter $p$, and $P_{1.1}$, $P_{0.9}$, and $P_{1}$ are the value of the measured variable when the model is run with parameter $p$ at 110%, 90%, and 100% of its value respectively.

* All parameters in Table @tbl-params were tested across all farms (all temperatures)
* Used a population of 100 in each farm

sens_data %>% 
  ggplot(aes(x = adj_param, y = mean_sens, ymin = mean_sens - sd_sens, ymax = mean_sens + sd_sens, fill = measure)) +
  geom_col(position = position_dodge(), width = 0.95, colour = "black", alpha = 0.75) +
  # geom_errorbar(position = position_dodge(width = 0.95), width = 0.3) +
  geom_hline(aes(yintercept = 0), linetype = "dashed") +
  scale_y_continuous(breaks = seq(-30,30,5), limits = c(-10,30)) +
  scale_x_discrete(limits = rev(param_names$name), labels = rev(param_names$lab)) +
  scale_fill_manual(values = c(weight = "salmon", total_excr = "steelblue")) +
  # scale_fill_manual(values = c("weight" = "salmon", "uneat" = "steelblue", "excr" = "darkgreen")) +
  labs(x = "Adjusted parameter", y = "Sensitivity") +
  theme_classic() + 
  theme(legend.position = "none", 
        strip.text = element_blank(), 
        aspect.ratio = 0.8,
        text = element_text(size = 12),
        axis.title.y = element_blank()) +
  coord_flip()
```

@@

```{r feeds}
feed_pal <- c("past" = brewer.pal(9, "Set1")[1], "reference" = brewer.pal(9, "Set1")[2], "future" = brewer.pal(9, "Set1")[3])
feed_params <- file.path(output_species_data_path, "feed_params.qs") %>% qs::qread()

df <- feed_params[["reference"]]
contribP_1 <- df[["Proteins"]]$proportion*df[["Proteins"]]$macro
tP_1 <- sum(contribP_1)
digestP_1 <- sum(df[["Proteins"]]$digest * contribP_1)/tP_1

contribL_1 <- df[["Lipids"]]$proportion*df[["Lipids"]]$macro
tL_1 <- sum(contribL_1)
digestL <- sum(df[["Lipids"]]$digest * contribL_1)/tL_1

contribC_1 <- df[["Carbohydrates"]]$proportion*df[["Carbohydrates"]]$macro
tC_1 <- sum(contribC_1)
digestC_1 <- sum(df[["Carbohydrates"]]$digest * contribC_1)/tC_1

df <- feed_params[["past"]]
contribP_2 <- df[["Proteins"]]$proportion*df[["Proteins"]]$macro
tP_2 <- sum(contribP_2)
digestP_2 <- sum(df[["Proteins"]]$digest * contribP_2)/tP_2

contribL_2 <- df[["Lipids"]]$proportion*df[["Lipids"]]$macro
tL_2 <- sum(contribL_2)
digestL <- sum(df[["Lipids"]]$digest * contribL_2)/tL_2

contribC_2 <- df[["Carbohydrates"]]$proportion*df[["Carbohydrates"]]$macro
tC_2 <- sum(contribC_2)
digestC_2 <- sum(df[["Carbohydrates"]]$digest * contribC_2)/tC_2

df <- feed_params[["future"]]
contribP_3 <- df[["Proteins"]]$proportion*df[["Proteins"]]$macro
tP_3 <- sum(contribP_3)
digestP_3 <- sum(df[["Proteins"]]$digest * contribP_3)/tP_3

contribL_3 <- df[["Lipids"]]$proportion*df[["Lipids"]]$macro
tL_3 <- sum(contribL_3)
digestL <- sum(df[["Lipids"]]$digest * contribL_3)/tL_3

contribC_3 <- df[["Carbohydrates"]]$proportion*df[["Carbohydrates"]]$macro
tC_3 <- sum(contribC_3)
digestC_3 <- sum(df[["Carbohydrates"]]$digest * contribC_3)/tC_3
```

[Possibly replace this with a better summary of the ingredients that make up each feed, and report these stats in Results under Uneaten feed]{.aside}

* Feed came from [@cottrell_origins_nodate; @cottrell_global_2020]
* The reference feed is `r round(100*tP_1,1)`% protein, `r round(100*tL_1,1)`% lipid, `r round(100*tC_1,1)`% carbohydrate. 
* The past feed is `r round(100*tP_2,1)`% protein, `r round(100*tL_2,1)`% lipid, `r round(100*tC_2,1)`% carbohydrate. 
* The future feed is `r round(100*tP_3,1)`% protein, `r round(100*tL_3,1)`% lipid, `r round(100*tC_3,1)`% carbohydrate. 
* Within the reference feed, protein is overall `r round(100*digestP_1,1)`% digestible, lipid is overall `r round(100*digestL,1)`% digestible, and carbohydrate is overall `r round(100*digestC_1,1)`% digestible.
* Within the reference feed, protein is overall `r round(100*digestP_2,1)`% digestible, lipid is overall `r round(100*digestL,1)`% digestible, and carbohydrate is overall `r round(100*digestC_2,1)`% digestible.
* Within the reference feed, protein is overall `r round(100*digestP_3,1)`% digestible, lipid is overall `r round(100*digestL,1)`% digestible, and carbohydrate is overall `r round(100*digestC_3,1)`% digestible.

## Population and biomass estimates

>>>>

library(magrittr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ggh4x)
library(here)
library(ggpubr)
library(cowplot)
library(patchwork)
library(qs)
library(arrow)
library(targets)
library(sf)
library(terra)
library(rnaturalearth)
library(rnaturalearthdata)
library(rnaturalearthhires)
library(dtplyr)
library(furrr)
library(RColorBrewer)
library(purrr)
library(future)
library(readxl)
library(units)
library(conflicted)
library(lubridate)
library(RColorBrewer)
conflicts_prefer(dplyr::select(), dplyr::filter(), dplyr::intersect(), .quiet = T)
# This markdown uses TinyTex - install with tinytex::install_tinytex()

remove_unit("g_fish")
remove_unit("kg_fish")
remove_unit("t_fish")
install_unit("g_fish")
install_unit("kg_fish", "1000 g_fish")
install_unit("t_fish", "1000 kg_fish")
```

@@

species_params <- qread(species_params_file)
pop_params <- qread(pop_params_file)

prettyplot <- function() {
  theme_classic() +
    theme(legend.position = "none",
          text = element_text(family = "serif", size = 12, colour = "black"),
          axis.title.x = element_text(vjust = 0.5),
          axis.title.y = element_text(hjust = 0.5))
}

feed_pal <- c("past" = "#E41A1C", "reference" = "#377EB8", "future" = "#4DAF4A")  
cohort_pal <- c("1" = "#E41A1C", "2" = "#377EB8", "3" = "#4DAF4A", "all" = "black")
macro_pal <- c(
  "protein" = "darkred", "P" = "darkred", 
  "carbohydrates" = "steelblue", "C" = "steelblue", 
  "lipids" = "darkgoldenrod1", "L" = "darkgoldenrod1"
)
```

```{r custom fuctions}
#| code-summary: Function to create sf polygons from bounding boxes

create_boxes <- function(box_list, label_positions, offset_deg) {
  boxes <- list()
  labels <- list()

  for(i in seq_along(box_list)) {
    name <- names(box_list)[i]
    coords <- box_list[[i]]
    pos <- label_positions[i]

    # Create polygon from bounding box coordinates
    box_coords <- matrix(c(
      coords[1], coords[3],  # lonmin, latmin
      coords[2], coords[3],  # lonmax, latmin  
      coords[2], coords[4],  # lonmax, latmax
      coords[1], coords[4],  # lonmin, latmax
      coords[1], coords[3]   # close polygon
    ), ncol = 2, byrow = TRUE)

    # Create sf polygon
    poly <- st_polygon(list(box_coords))
    boxes[[i]] <- st_sfc(poly, crs = 4326)

    # Calculate label position based on specification
    label_coords <- get_label_position(coords, pos, offset_deg[i])
    label_point <- st_point(c(label_coords[1], label_coords[2]))
    labels[[i]] <- st_sfc(label_point, crs = 4326)
  }

  # Combine into sf objects
  boxes_sf <- st_sf(
    name = names(box_list),
    letter = LETTERS[1:length(box_list)],
    geometry = do.call(c, boxes)
  )

  labels_sf <- st_sf(
    name = names(box_list), 
    letter = LETTERS[1:length(box_list)],
    geometry = do.call(c, labels)
  )

  return(list(boxes = boxes_sf, labels = labels_sf))
}

# Helper function to calculate label positions
get_label_position <- function(coords, position, offset) {
  lonmin <- coords[1]
  lonmax <- coords[2]
  latmin <- coords[3]
  latmax <- coords[4]

  # Parse position string
  parts <- strsplit(position, "_")[[1]]
  vertical <- parts[1]    # "top" or "bottom"
  horizontal <- parts[2]  # "left" or "right"
  location <- parts[3]    # "inside" or "outside"

  # Base coordinates for corners
  if(vertical == "bottom") {base_lat <- latmin} else {base_lat <- latmax}
  if(horizontal == "right") {base_lon <- lonmax} else {base_lon <- lonmin}

  # Apply offset for outside positioning
  if(location == "outside") {
    if(horizontal == "left") {base_lon <- base_lon-offset} else {base_lon <- base_lon+offset}
    if(vertical == "top") {base_lat <- base_lat + offset} else {base_lat <- base_lat - offset}
  }

  return(c(base_lon, base_lat))
}

lim_robin <- function(lons = c(144.0, 149.5), lats = c(-39.75, -44.00)) {
  coords <- data.frame(lon = lons, lat = lats) %>%
    st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
    st_transform(crs = "+proj=robin") %>% 
    st_coordinates()
  list(
    xlims = range(coords[, "X"]),
    ylims = range(coords[, "Y"])
  )
}

```

This document was last updated at `r format(Sys.time(), "%H:%M")` on `r format(Sys.time(), "%d %b, %Y")`.

# Methods
### Feeding rate

df %>% 
  ggplot(aes(x = water_temp, y = ing_pot)) +
  geom_line(linewidth = 0.75) +
  geom_line(aes(y = food_prov), linewidth = 0.75, linetype = "dotted") +
  scale_y_continuous(breaks = seq(0,1.5,0.25)) +
  scale_x_continuous(breaks = seq(-5,30,5), limits = c(-5,26)) +
  labs(x = expression("Water temperature ("*degree*"C)"), y = "Relative food ingested/provided") +
  prettyplot() +
  theme(text = element_text(colour = "black", family = "serif", size = 12))
```

@@

worldmap <- ne_countries(scale = "medium", returnclass = "sf") %>% st_transform(crs = "+proj=robin")
graticules <- st_graticule(worldmap, lon = seq(-180, 180, 30), lat = seq(-90, 90, 30))

inset_boxes <- list(   # lonmin, lonmax, latmin, latmax
  CAN1 = c(-132, -122, 47.75, 54.25),
  CAN2 = c(-70, -54, 43, 48.5),
  EUR = c(-26, 30, 51, 72),
  CHI = c(-77.5, -62.5, -56, -25),
  AUS = c(144, 149.5, -44, -39.75)
)
labels_spec <- c(CAN1 = "bottom_left_outside", CAN2 = "bottom_right_outside", EUR = "bottom_left_outside", CHI = "top_left_outside", AUS = "bottom_left_outside")
labels_offset <- c(CAN1 = 1, CAN2 = 2, EUR = 1.75, CHI = 1.75, AUS = 3.5)

box_data <- create_boxes(box_list = inset_boxes, # see function create_boxes above
                         label_positions = unname(labels_spec),
                         offset_deg = unname(labels_offset))
boxes_robinson <- st_transform(box_data$boxes, crs = "+proj=robin")
labels_robinson <- st_transform(box_data$labels, crs = "+proj=robin")
farm_temp_robinson <- st_sf(farm_temp_means) %>% st_transform(crs = "+proj=robin")

p1 <- ggplot() +
  geom_sf(data = graticules, color = "gray80", size = 0.3) +
  geom_sf(data = worldmap, fill = "white", color = "dimgray") +
  geom_sf(data = farm_temp_robinson, aes(color = mean_temp), size = 1) +
  coord_sf() +
  scale_color_viridis_c(option = "turbo") +
  guides(col = guide_colourbar(
    title = "Mean temperature (°C)",
    direction = "horizontal", position = "bottom",
    label.position = "bottom", title.position = "top", 
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 2,
    barwidth = 25,barheight = 1.5
  )) +
  geom_sf(data = boxes_robinson, fill = NA, color = "darkred", size = 1.65) +
  geom_sf_text(data = labels_robinson, aes(label = letter), 
               color = "darkred", size = 5, fontface = "bold", 
               hjust = 0.5, vjust = 0.5) +
  theme_void()

p2 <- farm_temp_means %>% 
  ggplot(aes(y = mean_temp, x = abs(lat), colour = country)) +
  geom_point(size = 2) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = expression("Absolute latitude ("*degree*")"), y = expression("Mean temperature ("*degree*"C)")) +
  prettyplot() +
  theme(legend.position = "top", legend.title = element_blank())

p3 <- farm_temp_means %>% 
  mutate(hemisphere = case_when(lat < 0 ~ "S", T ~ "N")) %>% 
  ggplot(aes(x = mean_temp, fill = hemisphere)) +
  geom_histogram(position = "identity", colour = "black", alpha = 0.5, binwidth = 0.5) +
  labs(x = expression("Mean temperature ("*degree*"C)"), y = "Frequency") +
  prettyplot() +
  theme(legend.position = "top", legend.title = element_blank())

@@

```{r patchwork-map-mercator}
#| fig-cap: |
#|   Example of a patchwork
#| fig-height: 9
#| fig-width: 9.5

no_margins <- function() {
  theme(legend.position = "none", 
        plot.margin = margin(0, 0, 0, 0),
        axis.title = element_blank())
}

inset_boxes_sm <- list(   # lonmin, lonmax, latmin, latmax
  CAN1 = c(-132, -122, 48.5, 54),
  CAN2 = c(-70, -54, 43, 48.5),
  EUR = c(-23, 28.5, 52, 71),
  CHI = c(-78, -62, -55.5, -27),
  AUS = c(144, 149, -43.75, -40.75)
)

inset_boxes_sm <- map(inset_boxes_sm, function(bx) {
  list(
    xlims = bx[1:2],
    ylims = bx[3:4],
    labx = bx[1],
    laby = bx[4]
    )
})

worldmap <- ne_countries(scale = "large", returnclass = "sf")
p1 <- ggplot() +
  geom_sf(data = worldmap, fill = "white", color = "dimgray") +
  geom_sf(data = farm_temp_means, aes(color = mean_temp, geometry = geometry), size = 1) +
  coord_sf() +
  scale_color_viridis_c(option = "turbo") +
  guides(col = guide_colourbar(
    title = "Mean temperature (°C)",
    direction = "vertical", position = "right",
    label.position = "right", title.position = "top", 
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 2,
    barwidth = 1.5, barheight = 20
  )) +
  prettyplot() +
  labs(y = "Latitude", x = "Longitude")

p_Can1 <- p1 + 
  coord_sf(xlim = inset_boxes_sm[["CAN1"]][["xlims"]], 
           ylim = inset_boxes_sm[["CAN1"]][["ylims"]])  +
  draw_label("A", size = 14, fontface = "bold", 
             x = inset_boxes_sm[["CAN1"]][["labx"]], 
             y = inset_boxes_sm[["CAN1"]][["laby"]], 
             hjust = -1, vjust = 0.5)
p_Can2 <- p1 + 
  coord_sf(xlim = inset_boxes_sm[["CAN2"]][["xlims"]], 
           ylim = inset_boxes_sm[["CAN2"]][["ylims"]])  +
  draw_label("B", size = 14, fontface = "bold", 
             x = inset_boxes_sm[["CAN2"]][["labx"]], 
             y = inset_boxes_sm[["CAN2"]][["laby"]], 
             hjust = 0.5, vjust = 0)
p_Eur <- p1 + 
  coord_sf(xlim = inset_boxes_sm[["EUR"]][["xlims"]], 
           ylim = inset_boxes_sm[["EUR"]][["ylims"]])  +
  draw_label("C", size = 14, fontface = "bold", 
             x = inset_boxes_sm[["EUR"]][["labx"]], 
             y = inset_boxes_sm[["EUR"]][["laby"]], 
             hjust = -1, vjust = 0.5)
p_Chi <- p1 + 
  coord_sf(xlim = inset_boxes_sm[["CHI"]][["xlims"]], 
           ylim = inset_boxes_sm[["CHI"]][["ylims"]])  +
  theme(legend.position = "right") +
  draw_label("D", size = 14, fontface = "bold", 
             x = inset_boxes_sm[["CHI"]][["labx"]], 
             y = inset_boxes_sm[["CHI"]][["laby"]], 
             hjust = 0, vjust = 0)
p_Aus <- p1 + 
  coord_sf(xlim = inset_boxes_sm[["AUS"]][["xlims"]], 
           ylim = inset_boxes_sm[["AUS"]][["ylims"]])  +
  draw_label("E", size = 14, fontface = "bold", 
             x = inset_boxes_sm[["AUS"]][["labx"]], 
             y = inset_boxes_sm[["AUS"]][["laby"]], 
             hjust = 0, vjust = 0)

grid_1 <- cowplot::plot_grid(p_Can1, p_Can2, ncol = 2, rel_widths =  c(1, 1.35))
grid_2 <- cowplot::plot_grid(p_Eur,  p_Aus,  nrow = 2, rel_heights = c(1, 1))
grid_3 <- cowplot::plot_grid(grid_2, p_Chi,  ncol = 2, rel_widths =  c(1, 1))

cowplot::plot_grid(
  grid_1, grid_3,
  nrow = 2,
  rel_heights = c(1, 1.5)        # relative row heights
)
```

@@

Parameter sensitivity was tested by varying each parameter in @tbl-params by $\pm 10$% and calculating sensitivity as $S_p = \frac{(P_{1.1} - P_{0.9})}{0.2 \times P_1}$, where $S$ is the sensitivity of the measured variable to changes in parameter $p$, and $P_{1.1}$, $P_{0.9}$, and $P_{1}$ are the value of the measured variable when the model is run with parameter $p$ at 110%, 90%, and 100% of its value respectively. 
All parameters in Table @tbl-params were tested. The population-level parameters ($\overline{W}$, $\Delta{W}$, $\overline{I_{max}}$, $\Delta{I_{max}}$, $\overline{F}$, $\Delta{F}$) were tested with a population of of 5000 on a random 10% of farms, while all other parameters were tested across all farms with a population of 10.

@@

```{r feeds}
feed_params <- file.path(output_species_data_path, "feed_params.qs") %>%
  qs::qread()

feed_summary <- map_dfr(names(feed_params), function(feed) {
  df <- feed_params[[feed]]
  map_dfr(c("Proteins", "Lipids", "Carbohydrates"), function(macro) {
    contrib <- df[[macro]]$proportion * df[[macro]]$macro
    total <- sum(contrib)
    digest <- sum(df[[macro]]$digest * contrib) / total
        tibble(
      feed = feed,
      macro = macro,
      total = total,
      digest = digest
    )
  })
}) %>%
  mutate(macro = factor(macro, levels = c("Proteins", "Lipids", "Carbohydrates"), labels = c("P", "L", "C")),
         feed = factor(feed, levels = c("reference", "past", "future")),
         total_r = total * 100 %>% round(1),
         digest_r = digest * 100 %>% round(1))
```

Feeds came from [@cottrell_origins_nodate; @cottrell_global_2020].

[Possibly replace this with a better summary of the ingredients that make up each feed, and report these stats in Results under Uneaten feed?]{.aside}

The reference feed was 
`r round(100*filter(feed_summary, feed == "reference", macro == "P")$total, 1)`% protein, 
`r round(100*filter(feed_summary, feed == "reference", macro == "L")$total, 1)`% lipid, and 
`r round(100*filter(feed_summary, feed == "reference", macro == "C")$total, 1)`% carbohydrate, the past feed was 
`r round(100*filter(feed_summary, feed == "past", macro == "P")$total, 1)`% protein, 
`r round(100*filter(feed_summary, feed == "past", macro == "L")$total, 1)`% lipid, and 
`r round(100*filter(feed_summary, feed == "past", macro == "C")$total, 1)`% carbohydrate, and the future feed was 
`r round(100*filter(feed_summary, feed == "future", macro == "P")$total, 1)`% protein, 
`r round(100*filter(feed_summary, feed == "future", macro == "L")$total, 1)`% lipid, and 
`r round(100*filter(feed_summary, feed == "future", macro == "C")$total, 1)`% carbohydrate.

Ingredients are tracked seperately through the model with their own digestibility coefficients. Overall, proteins within the reference feed were 
`r round(100*filter(feed_summary, feed == "reference", macro == "P")$digest, 1)`% digestible, proteins within the past feed were 
`r round(100*filter(feed_summary, feed == "past", macro == "P")$digest, 1)`% digestible, and proteins within the future feed were 
`r round(100*filter(feed_summary, feed == "future", macro == "P")$digest, 1)`% digestible.
The digestibility of the other macronutrients are assumed to remain constant across all ingredients at `r round(100*filter(feed_summary, feed == "past", macro == "C")$digest, 1)`% for carbohydrates and `r round(100*filter(feed_summary, feed == "past", macro == "L")$digest, 1)`% for lipids.

```{r fig-feed-summary-ideas}
#| fig-height: 7.5
#| fig-cap: | 
#|   Ideas for how to present the differences between feeds. Top plot shows the protein (red), lipids (yellow) and carbohydrate (blue) composition, with the digestible portion shown darker and the undigestible portion shown lighter. Bottom plot shows the proportion of each feed composed of each types of ingredient.

p1 <- feed_summary %>%
  mutate(digest_total = digest * total) %>% 
  ggplot(aes(x = feed, y = total, fill = macro)) +
  geom_col(position = "dodge", alpha = 0.55, colour = "black") +
  geom_col(mapping = aes(y = digest_total), position = "dodge", alpha = 1, colour = "black") +
  scale_fill_manual(values = macro_pal) +
  scale_y_continuous(breaks = seq(0,0.5,0.1), labels = seq(0,50,10)) +
  scale_x_discrete(labels = c("reference" = "Reference", "past" = "Past", "future" = "Future")) +
  labs(x = "Feed", y = "Proportion (%)") +
  prettyplot()

feed_ings <- rbind(
  feed_params[[1]][[1]] %>% mutate(feed = as.factor("reference")), 
  feed_params[[2]][[1]] %>% mutate(feed = as.factor("past")), 
  feed_params[[3]][[1]] %>% mutate(feed = as.factor("future"))
) %>%
  rename(protein = macro) %>%
  mutate(
    category = case_when(
      str_detect(ingredient, "fish-oil") ~ "fish",
      str_detect(ingredient, "fishmeal") ~ "fish",
      str_detect(ingredient, "krill") ~ "other-marine",
      str_detect(ingredient, "soy") ~ "legume",
      str_detect(ingredient, "pea") ~ "legume",
      str_detect(ingredient, "guar") ~ "legume",
      str_detect(ingredient, "faba") ~ "legume",
      str_detect(ingredient, "corn") ~ "wheat-corn",
      str_detect(ingredient, "wheat") ~ "wheat-corn",
      str_detect(ingredient, "chicken") ~ "chicken",
      str_detect(ingredient, "bacteri") ~ "bact-fungi-micro",
      str_detect(ingredient, "yeast") ~ "bact-fungi-micro",
      str_detect(ingredient, "spirulina") ~ "bact-fungi-micro",
      str_detect(ingredient, "bsf") ~ "insect",
      T ~ "other-plant"
    ) %>% 
      factor(levels = c("fish", "other-marine", "chicken", "wheat-corn", "legume", "other-plant", "insect", "bact-fungi-micro"))
  )
